---
title: 'Loading Workflow Objects'
teaching: 10
exercises: 2
---

```{r}
#| label: setup
#| echo: FALSE
#| message: FALSE
#| warning: FALSE

if (interactive()) {
  setwd("episodes")
}

library(targets)
source("files/lesson_functions.R")
```

:::::::::::::::::::::::::::::::::::::: questions 

- Where does the workflow happen?
- How can we inspect the objects built by the workflow?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Explain where `targets` runs the workflow and why
- Be able to load objects built by the workflow into your R session

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: instructor

Episode summary: Show how to get at the objects that we built

:::::::::::::::::::::::::::::::::::::

## Where does the workflow happen?

So we just finished running our workflow.
Now you probably want to look at its output.
But, if we just call the name of the object (for example, `penguins_data`), we get an error.
```{r}
#| label: error
penguins_data
```

Where are the results of our workflow?

::::::::::::::::::::::::::::::::::::: instructor

- To reinforce the concept of `targets` running in a separate R session, you may want to pretend trying to run `penguins_data`, then feigning surprise when it doesn't work and using it as a teaching moment (errors are pedagogy!).

::::::::::::::::::::::::::::::::::::::::::::::::

We don't see the workflow results because `targets` **runs the workflow in a separate R session** that we can't interact with.
This is for reproducibility---the objects built by the workflow should only depend on the code in your project, not any commands you may have interactively given to R.

Fortunately, `targets` has a function that can load objects built by the workflow into our current session, `tar_load()`.
Let's see how this works.

## tar_load()

`tar_load()` loads an object built by the workflow into the current session.
Its first argument is the name of the object you want to load.
Let's use this to load `penguins_data` and get an overview of the data with `summary()`.

```{r}
#| label: targets-run-hide
#| echo: FALSE
# When building the Rmd, each instance of the workflow is isolated, so need
# to re-run
plan_1_dir <- make_tempdir()
pushd(plan_1_dir)
write_example_plan("plan_1.R")
tar_make(reporter = "silent")
penguins_csv_file_hide <- tar_read(penguins_csv_file)
penguins_data_hide <- tar_read(penguins_data)
popd()
```

```{r}
#| label: targets-load
#| echo: [2, 3]
pushd(plan_1_dir)
tar_load(penguins_data)
summary(penguins_data)
popd()
```

Note that `tar_load()` is used for its **side-effect**---loading the desired object into the current R session.
It doesn't actually return a value.

If you want to learn more about an alternate method to inspect objects built by the pipeline, [see the Appendix](extras.Rmd#an-alternate-way-to-inspect-target-objects).

## The targets cache

If you close your R session, then re-start it and use `tar_load()` or `tar_read()`, you will notice that it can still load the workflow objects.
In other words, the workflow output is **saved across R sessions**.
How is this possible?

You may have noticed a new folder has appeared in your project, called `_targets`.
This is the **targets cache**.
It contains all of the workflow output; that is how we can load the targets built by the workflow even after quitting then restarting R.

**You should never edit the contents of the cache by hand**.\*
Doing so would make your analysis non-reproducible!

Note that it is possible to "reset" your workflow by simply deleting the entire `_targets` folder.
Of course, this means you will need to run everything over again, so don't do this lightly!

\*: There is one exception, [described in the Appendix](extras.Rmd#storing-user-data-files-in-the-cache).

::::::::::::::::::::::::::::::::::::: keypoints 

- `targets` workflows are run in a separate, non-interactive R session
- `tar_load()` loads a workflow object into the current R session
- The `_targets` folder is the cache and generally should not be edited by hand

::::::::::::::::::::::::::::::::::::::::::::::::
