---
title: 'Deploying Targets on HPC'
teaching: 10
exercises: 2
---

```{R}
#| echo: false
# Exit sensibly when Slurm isn't installed 
if (!nzchar(Sys.which("sbatch"))){
  knitr::knit_exit("sbatch was not detected. Likely Slurm is not installed. Exiting.")
}
```

::: questions 

- Why would we use HPC to run Targets workflows?
- How can we run Targets workflows on Slurm?

:::

::: objectives

- Be able to run a Targets workflow on Slurm
- Understand how workers relate to targets 
- Know how to configure Slurm jobs within targets
- Be able to create a pipeline with heterogeneous workers

:::

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false
library(targets)
library(tarchetypes)
library(quarto) # don't actually need to load, but put here so renv catches it
source("files/lesson_functions.R") # nolint

# Increase width for printing tibbles
options(width = 140)
```

## Advantages of HPC

If your analysis involves computationally intensive or long-running tasks such as training machine learning models or processing very large amounts of data, it will quickly become infeasible to use a single machine to run this.
If you have access to a High Performance Computing (HPC) cluster, you can leverage the numerous machines with Targets to scale up your analysis.
This differs from the exucution we have learned so far, which spawns extra R processes on the *same machine* to speed up execution.

## Configuring Targets for Slurm

::: {.prereq}

### Install required packages

You will need to install `crew.cluster` to enable the HPC integration:

```{r}
#| label: 'install-crew.cluster'
#| eval: false
install.packages("crew.cluster")
```
:::

To adapt Targets to use the Slurm HPC scheduler, we change the `controller`.
In this section we will assume that our HPC uses Slurm as its job scheduler, but you can use other schedulers such as PBS/TORQUE, Sun Grid Engine (SGE) or LSF.

In the Parallel Processing section, we used the following configuration:
```{R}
#| label: one-machine-crew
tar_option_set(
  controller = crew_controller_local(workers = 2)
)
```
To configure this for Slurm, we swap out the controller with [`crew_controller_slurm()`](https://wlandau.github.io/crew.cluster/reference/crew_controller_slurm.html) a new one from the [`crew.cluster`](https://wlandau.github.io/crew.cluster/index.html) package:

```{R}
#| label = "slurm-crew",
#| eval = FALSE,
#| code = readLines("files/plans/plan_slurm.R")[11:17]
```

::: callout
If you were using a scheduler other than Slurm, you would instead select [`crew_controller_lsf()`](https://wlandau.github.io/crew.cluster/reference/crew_controller_lsf.html), [`crew_controller_pbs()`](https://wlandau.github.io/crew.cluster/reference/crew_controller_pbs.html) or [`crew_controller_sge()`](https://wlandau.github.io/crew.cluster/reference/crew_controller_sge.html) instead of `crew_controller_slurm()`.

These functions have their own unique arguments which are associated with the scheduler.
:::

There are a number of options you can pass to `crew_controller_slurm()` to fine-tune the Slurm execution, [which you can find here](https://wlandau.github.io/crew.cluster/reference/crew_controller_slurm.html).
Here we are only using three:

  * `workers` sets the number of jobs that are submitted to Slurm to process targets.
  * `script_lines` adds some lines to the Slurm submit script used by Targets. This is useful for loading Environment Modules as we have done here.
  * `slurm_memory_gigabytes_per_cpu` specifies the amount of memory we need.

Let's run the modified workflow:

```{R, eval=FALSE}
#| label = "slurm-workflow-show",
#| eval = FALSE,
#| code = readLines("files/plans/plan_slurm.R")
```
```{R}
#| label: slurm-workflow-hide
#| echo: false
tar_dir({
  write_example_plan("plan_slurm.R")
  tar_make()
})
```

We've successfully transferred our analysis onto a Slurm cluster!

::: challenge
## Increasing Resources

Q: How would you modify your `_targets.R` if your functions needed 2 CPUs?

::: hint
Check the arguments for [`crew_controller_slurm`](https://wlandau.github.io/crew.cluster/reference/crew_controller_slurm.html#arguments-1).
:::
::: solution
```R
tar_option_set(
  controller = crew_controller_slurm(
    workers = 3,
    script_lines = "module load R",
    slurm_memory_gigabytes_per_cpu = 1,
    # Added this
    slurm_cpus_per_task = 2
  )
)
```
:::
:::

## SBATCH Options

The `script_lines` argument shown above can also be used to add `#SBATCH` flags, to configure your worker job.
Each entry in the vector will be treated as a new line to be added to the `sbatch` script that is generated.
However, you have to be careful to put all of your `#SBATCH` lines before any other bash commands.
`sbatch` flags [are listed here](https://slurm.schedmd.com/sbatch.html#SECTION_OPTIONS) in the Slurm documentation.
For instance, to request that the worker has a GPU available, you could do the following:
```R
tar_option_set(
  controller = crew_controller_slurm(
    workers = 3,
    script_lines = c(
      "#SBATCH --gres=gpu:1",
      "module load R"
    ),
    slurm_memory_gigabytes_per_cpu = 1
  ) 
)
```

In general, it's better to use a dedicated `crew_controller_slurm` argument than to use `script_lines`, if one exists.
For example, prefer `slurm_cpus_per_task=2` to `script_lines="--cpus-per-task=2"` and set `name="my_name"` rather than using `script_lines="--job-name=my_name"`.

## HPC Workers

`crew` uses a persistent worker strategy.
This means that `crew` does not submit one Slurm job for each target. 
Instead, you define a pool of workers when configuring the workflow.
In our example above we specified a maximum of 3 workers.
For each worker, `crew` submits a single Slurm job, and these workers will process multiple targets over their lifetime.

We can verify that this has happened using `sacct`, which we can use to query information about our past jobs.
All the Slurm jobs with the same hash (the part after `crew-`) belong to the same Slurm controller:

```{bash}
sacct --starttime now-5minutes --allocations
```

The upside of this approach is that we don't have to know how long each target takes to build, or what resources it needs.
It also means that we don't submit a lot of jobs, making our Slurm usage easy to monitor.

The downside of this mechanism is that **the resources of the worker have to be sufficient to build all of your targets**. 
In other words, you need to work out the maximum RAM and CPUs used across all of your targets, and specify those maximum resources in the `crew_controller_slurm()` function.

::: challenge
## Choosing a Worker

Q: Say we have two targets. One uses 100 GB of RAM and 1 CPU, and the other needs 10 GB of RAM and 8 CPUs to run a multi-threaded function. What worker configuration do we use?

::: solution
We need to choose the maximum of all resources if we have a single worker.
It will need 100 GB of RAM and 8 CPUs.
To do this we might use a controller a bit like this:
```{R, results="hide"}
crew_controller_slurm(
  name = "cpu_worker",
  workers = 3,
  script_lines = "module load R",
  slurm_cpus_per_task = 8,
  slurm_memory_gigabytes_per_cpu = 100 / 8
)
```
:::
:::

## Heterogeneous Workers

In some cases we may prefer to use more than one different Slurm job processing our targets, especially if some of our targets need different hardware from others, such as a GPU.
When we do this, we say we have "heterogeneous workers", meaning that not all worker jobs are the same as each other.
To do this, we firstly define each worker configuration by adding the `name` argument to `crew_controller_slurm`:

```{R}
#| eval = FALSE,
#| code = readLines("files/plans/plan_slurm_memory.R")[8:19]
```

Next, we tell Targets about these controllers using `tar_option_set` as before, with one difference: we have to combine them in a controller group:

```{R}
#| eval = FALSE,
#| code = readLines("files/plans/plan_slurm_memory.R")[20:23]
tar_option_set(
  controller = crew_controller_group(small_memory, big_memory)
)
```

Then we specify each controller by name in each target definition:

```{R}
#| eval = FALSE,
#| code = readLines("files/plans/plan_slurm_memory.R")[24:39]
```

When we run the pipeline, we can see the differing results:
```{R}
#| label: het-example-show
#| eval: false
tar_make()
tar_read(big_memory_task)
tar_read(small_memory_task)
```
```{R}
#| label: het-example-hide
#| eval: true
#| echo: false
tar_dir({
  write_example_plan("plan_slurm_memory.R")
  tar_make()
  tar_read(big_memory_task) |> print()
  tar_read(small_memory_task) |> print()
})
```

::: challenge
## Mixing GPU and CPU targets

Q: Say we have the following targets workflow. How would we modify it so that `gpu_task` is only run in a GPU Slurm job?
```{R, eval=FALSE}
graphics_devices <- function(){
  system2("lshw", c("-class", "display"), stdout=TRUE, stderr=FALSE)
}

tar_plan(
  tar_target(
    cpu_hardware,
    graphics_devices()
  ),
  tar_target(
    gpu_hardware,
    graphics_devices()
  )
)
```

::: hint
You will need to define two different crew controllers.
Also, [you will need to request a GPU from Slurm](https://slurm.schedmd.com/gres.html#Running_Jobs).
You can find an example of this above.
:::
::: solution
```{R}
#| label = 'heterogeneous-controllers-show',
#| eval = FALSE,
#| code = readLines("files/plans/plan_slurm_gpu.R")[13:35]
```
```{R}
#| label: heterogeneous-controllers-hide
#| eval: true
#| echo: false
tar_dir({
  write_example_plan("plan_slurm_gpu.R")
  tar_make()
  tar_load_everything()
})
```
```{R}
#| label: heterogeneous-controllers-cpu-show
#| eval: false
tar_read("cpu_hardware")
```
```{R}
#| label: heterogeneous-controllers-cpu-hide
#| eval: true
#| echo: false
cpu_hardware
```
```{R}
#| label: heterogeneous-controllers-gpu-show
#| eval: false
tar_read("gpu_hardware")
```
```{R}
#| label: heterogeneous-controllers-gpu-hide
#| eval: true
#| echo: false
gpu_hardware
```
:::
:::

::::::::::::::::::::::::::::::::::::: keypoints 

- `crew.cluster::crew_controller_slurm()` is used to configure a workflow to use Slurm
- Crew uses persistent workers on HPC, and you need to choose your resources accordingly
- You can create heterogeneous workers by using multiple calls to `crew_controller_slurm(name=)`

::::::::::::::::::::::::::::::::::::::::::::::::
