---
title: 'Parallel Processing'
teaching: 10
exercises: 2
---

:::::::::::::::::::::::::::::::::::::: questions 

- How can we build targets in parallel?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Be able to build targets in parallel

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: instructor

Episode summary: Show how to use parallel processing

:::::::::::::::::::::::::::::::::::::

```{r}
#| label: setup
#| echo: FALSE
#| message: FALSE
#| warning: FALSE
library(targets)
library(tarchetypes)
library(broom)
source("https://raw.githubusercontent.com/joelnitta/targets-workshop/main/episodes/files/functions.R?token=$(date%20+%s)") # nolint

# Increase width for printing tibbles
options(width = 140)
```

Once a pipeline starts to include many targets, you may want to think about parallel processing.
This takes advantage of multiple processors in your computer to build multiple targets at the same time.

::::::::::::::::::::::::::::::::::::: {.callout}

## When to use parallel processing

Parallel processing should only be used if your workflow has a structure such that it makes sense---if your workflow only consists of a linear sequence of targets, then there is nothing to parallelize.

:::::::::::::::::::::::::::::::::::::

`targets` includes support for high-performance computing, cloud computing, and various parallel backends.
Here, we assume you are running this analysis on a laptop and so will use a relatively simple backend.
If you are interested in high-performance computing, [see the `targets` manual](https://books.ropensci.org/targets/hpc.html).

### Install R packages for parallel computing

For this demo, we will use the [`future` backend](https://github.com/HenrikBengtsson/future).

::::::::::::::::::::::::::::::::::::: {.prereq}

### Install required packages

You will need to install several packages to use the `future` backend:

```{r}
#| label: install-future
#| eval: false
install.packages("future")
install.packages("future.callr")
```

:::::::::::::::::::::::::::::::::::::

### Set up workflow

There are a few things you need to change to enable parallel processing with `future`:

- Load the `future` and `future.callr` packages
- Add a line with `plan(callr)`
- When you run the pipeline, use `tar_make_future(workers = 2)` instead of `tar_make()`

Here, `workers = 2` is the number of processes to run in parallel. You may increase this up to the number of cores available on your machine.

Make these changes to the penguins analysis. It should now look like this:

```{r}
#| label: example-model-show-setup
#| eval: FALSE
source("R/packages.R")
source("R/functions.R")

plan(callr)

tar_plan(
  # Load raw data
  tar_file_read(
    penguins_data_raw,
    path_to_file("penguins_raw.csv"),
    read_csv(!!.x, show_col_types = FALSE)
  ),
  # Clean data
  penguins_data = clean_penguin_data(penguins_data_raw),
  # Build models
  models = list(
    combined_model = lm(
      bill_depth_mm ~ bill_length_mm, data = penguins_data),
    species_model = lm(
      bill_depth_mm ~ bill_length_mm + species, data = penguins_data),
    interaction_model = lm(
      bill_depth_mm ~ bill_length_mm * species, data = penguins_data)
  ),
  # Get model summaries
  tar_target(
    model_summaries,
    glance_with_mod_name(models),
    pattern = map(models)
  ),
  # Get model predictions
  tar_target(
    model_predictions,
    augment_with_mod_name(models),
    pattern = map(models)
  )
)
```

There is still one more thing we need to modify only for the purposes of this demo: if we ran the analysis in parallel now, you wouldn't notice any difference in compute time because the functions are so fast.

So let's make "slow" versions of `glance_with_mod_name()` and `augment_with_mod_name()` using the `Sys.sleep()` function, which just tells the computer to wait some number of seconds.
This will simulate a long-running computation and enable us to see the difference between running sequentially and in parallel.

Add these functions to `functions.R` (you can copy-paste the original ones, then modify them):

```{r}
#| label: slow-funcs
#| eval: false

glance_with_mod_name_slow <- function(model_in_list) {
  Sys.sleep(4)
  model_name <- names(model_in_list)
  model <- model_in_list[[1]]
  broom::glance(model) |>
    mutate(model_name = model_name)
}

augment_with_mod_name_slow <- function(model_in_list) {
  Sys.sleep(4)
  model_name <- names(model_in_list)
  model <- model_in_list[[1]]
  broom::augment(model) |>
    mutate(model_name = model_name)
}
```

Then, change the plan to use the "slow" version of the functions:

```{r}
#| label: example-model-show-9
#| eval: FALSE
source("R/packages.R")
source("R/functions.R")

plan(callr)

tar_plan(
  # Load raw data
  tar_file_read(
    penguins_data_raw,
    path_to_file("penguins_raw.csv"),
    read_csv(!!.x, show_col_types = FALSE)
  ),
  # Clean data
  penguins_data = clean_penguin_data(penguins_data_raw),
  # Build models
  models = list(
    combined_model = lm(
      bill_depth_mm ~ bill_length_mm, data = penguins_data),
    species_model = lm(
      bill_depth_mm ~ bill_length_mm + species, data = penguins_data),
    interaction_model = lm(
      bill_depth_mm ~ bill_length_mm * species, data = penguins_data)
  ),
  # Get model summaries
  tar_target(
    model_summaries,
    glance_with_mod_name_slow(models),
    pattern = map(models)
  ),
  # Get model predictions
  tar_target(
    model_predictions,
    augment_with_mod_name_slow(models),
    pattern = map(models)
  )
)
```

Finally, run the pipeline with `tar_make_future()`.

```{r}
#| label: example-model-hide-9
#| echo: FALSE
tar_dir({
  # Non-slow functions
  write_example_plan(7)
  tar_make(reporter = "silent")
  # Slow functions
  write_example_plan(8)
  tar_make_future(workers = 2)
})
```

Notice that although the time required to build each individual target is about 4 seconds, the total time to run the entire workflow is less than the sum of the individual target times! That is proof that processes are running in parallel **and saving you time**.

The unique and powerful thing about targets is that **we did not need to change our custom function to run it in parallel**. We only adjusted *the workflow*. This means it is relatively easy to refactor (modify) a workflow for running sequentially locally or running in parallel in a high-performance context.

Now that we have demonstrated how this works, you can change your analysis plan back to the original versions of the functions you wrote.

::::::::::::::::::::::::::::::::::::: keypoints 

- Dynamic branching creates multiple targets with a single command
- You usually need to write custom functions so that the output of the branches includes necessary metadata 
- Parallel computing works at the level of the workflow, not the function

::::::::::::::::::::::::::::::::::::::::::::::::
